---
title: "Capstone Project"
output: html_notebook
---
<center>  <h3> Group 9 </h2> </center>
---

### Include required libraries

```{r}
# For decision tree
install.packages("party")
library("sandwich")
library("party")

# For colorpanel
install.packages("gplots")
library("gplots")

# For KNN Model
install.packages('class')
library("class")
install.packages("gmodels")
library("gdata")
library("gmodels")


# For plot confusion matrix
install.packages("cvms")
library(cvms)
library(tibble)

# For random Forest
install.packages("randomForest")
library(randomForest)

# For Feature selection
install.packages("MASS")
install.packages("leaps")
library(MASS) # for stepwise regression like forward selection and backward deletion
library(leaps) # for subsetting or all subset regression
```

### Include CSV file

```{r}
heart_dataset <- read.csv("heart.csv", stringsAsFactors = FALSE,header = TRUE)
heart_dataset
View(heart_dataset)
```

### Check Data types

```{r}
sapply(heart_dataset,class)
```

### View Summary of Dataset

```{r}
summary(heart_dataset)
```

### Get First 6 rows data

```{r}
head(heart_dataset)
```

### Check sum of null values

```{r}
sum(is.na(heart_dataset))
```
```{r}
which(is.na(heart_dataset))
```

### BoxPlot

```{r}
boxplot(heart_dataset$Age,horizontal = TRUE)
boxplot(heart_dataset$RestingBP,horizontal = TRUE)
boxplot(heart_dataset$Cholesterol,horizontal = TRUE)
boxplot(heart_dataset$MaxHR,horizontal = TRUE)
boxplot(heart_dataset$Oldpeak,horizontal = TRUE)
```
> Remove outliers of RestingBP

```{r}
nrow(heart_dataset)
Q1_RestingBP <- quantile(heart_dataset$RestingBP, .25)
Q3_RestingBP <- quantile(heart_dataset$RestingBP, .75)
IQR_RestingBP <- IQR(heart_dataset$RestingBP)

heart_dataset$Outlier <- (heart_dataset$RestingBP < (Q1_RestingBP - 1.5*IQR_RestingBP) | heart_dataset$RestingBP > (Q3_RestingBP + 1.5*IQR_RestingBP))

avg = mean(heart_dataset$RestingBP[!heart_dataset$Outlier])

heart_dataset$RestingBP[heart_dataset$Outlier] = avg

#heart_dataset <- subset(heart_dataset, heart_dataset$RestingBP > (Q1_RestingBP - 1.5*IQR_RestingBP) & heart_dataset$RestingBP< (Q3_RestingBP + 1.5*IQR_RestingBP))

boxplot(heart_dataset$RestingBP,horizontal = TRUE)
```

> Remove outliers of Cholesterol

```{r}
nrow(heart_dataset)
Q1_Cholesterol <- quantile(heart_dataset$Cholesterol, .25)
Q3_Cholesterol <- quantile(heart_dataset$Cholesterol, .75)
IQR_Cholesterol <- IQR(heart_dataset$Cholesterol)
IQR_Cholesterol

#heart_dataset$Outlier1 <- (heart_dataset$Cholesterol < (Q1_Cholesterol - 1.5*IQR_Cholesterol) | heart_dataset$Cholesterol > (Q3_Cholesterol + 1.5*IQR_Cholesterol))

#avg1 = mean(heart_dataset$Cholesterol[!heart_dataset$Outlier1])

#heart_dataset$Cholesterol[heart_dataset$Outlier1] = avg1

heart_dataset <- subset(heart_dataset, heart_dataset$Cholesterol> (Q1_Cholesterol - 1.5*IQR_Cholesterol) & heart_dataset$Cholesterol< (Q3_Cholesterol + 1.5*IQR_Cholesterol))
#boxplot(heart_dataset$Cholesterol,horizontal = TRUE)
nrow(heart_dataset)
```

> Remove outliers of MaxHR

```{r}
Q1_MaxHR <- quantile(heart_dataset$MaxHR, .25)
Q3_MaxHR <- quantile(heart_dataset$MaxHR, .75)
IQR_MaxHR <- IQR(heart_dataset$MaxHR)
IQR_MaxHR
#heart_dataset <- subset(heart_dataset, heart_dataset$MaxHR> (Q1_MaxHR - 1.5*IQR_MaxHR) & heart_dataset$MaxHR< (Q3_MaxHR + 1.5*IQR_MaxHR))
heart_dataset$Outlier <- (heart_dataset$MaxHR < (Q1_MaxHR - 1.5*IQR_MaxHR) | heart_dataset$MaxHR > (Q3_MaxHR + 1.5*IQR_MaxHR))

avg = mean(heart_dataset$MaxHR[!heart_dataset$Outlier])

heart_dataset$MaxHR[heart_dataset$Outlier] = avg

boxplot(heart_dataset$MaxHR,horizontal = TRUE)

```



```{r}
Q1_Oldpeak <- quantile(heart_dataset$Oldpeak, .25)
Q3_Oldpeak <- quantile(heart_dataset$Oldpeak, .75)
IQR_Oldpeak <- IQR(heart_dataset$Oldpeak)
IQR_Oldpeak
#heart_dataset <- subset(heart_dataset, heart_dataset$Oldpeak> (Q1_Oldpeak - 1.5*IQR_Oldpeak) & heart_dataset$Oldpeak< (Q3_Oldpeak + 1.5*IQR_Oldpeak))

heart_dataset$Outlier <- (heart_dataset$Oldpeak < (Q1_Oldpeak - 1.5*IQR_Oldpeak) | heart_dataset$Oldpeak > (Q3_Oldpeak + 1.5*IQR_Oldpeak))

avg = mean(heart_dataset$Oldpeak[!heart_dataset$Outlier])

heart_dataset$Oldpeak[heart_dataset$Outlier] = avg

boxplot(heart_dataset$Oldpeak,horizontal = TRUE)
nrow(heart_dataset)
```



#### Histogram of resting BP

```{r}
hist(heart_dataset$RestingBP,main="HIstogram to show RestingBP",xlab="Blood pressure In mm or Hg",xlim=c(80,180),col="skyblue")
```

#### Histogram to show Max Heart Rate

```{r}
hist(heart_dataset$MaxHR, main="HIstogram to show Max Heart Rate",xlab="Heart Rate Value",xlim=c(60,202),col=c("pink","grey","skyblue"))
```

#### Histogram to show Cholesterol

```{r}
hist(heart_dataset$Cholesterol, main="HIstogram to show Cholesterol Level",xlab="Cholesterol Value")
```


#### Heart Disease between meales and Females

> Heart Disease ratio between Males and Females 

```{r}
counts <- table(heart_dataset$HeartDisease, heart_dataset$Sex)
counts
barplot(counts, main="Haert Disease between Males and Females",
  xlab="Gender", col=c("grey","skyblue"),
  legend = rownames(counts), beside=TRUE, width = c(0.5,0.5))
```

#### Heart Disease based on Chest Pain Type 

 > Which type of chest pain cause heart disease
 
```{r}
counts_chestpain <- table(heart_dataset$HeartDisease, heart_dataset$ChestPainType)
counts_chestpain
barplot(counts_chestpain, main="Heart Disease based on Chest Pain Type",
  xlab="Chest Pain Type", col=c("pink","skyblue"),
  legend = rownames(counts_chestpain), beside=TRUE, width = c(0.5,0.5))
```

### Correlation

> Which attribute is more contributing in CVDs problem.

```{r}
subset <- heart_dataset [,c("HeartDisease", "RestingBP", "Cholesterol","FastingBS", "MaxHR", "Oldpeak")]
plot(subset)
cor(subset)
heatmap(cor(subset), col = redblue(100), margins=c(8,6))
```

### Classification

#### Linear regression

> is MaxHR is really playing a good role in RestingBP ?? 

```{r}
modelBP <- lm(RestingBP ~ MaxHR, data = subset)
summary(modelBP)
```
> is Cholesterol is really playing a good role in RestingBP ??

```{r}
Ch_modelBP <- lm(RestingBP ~ Cholesterol, data = subset)
summary(Ch_modelBP)
```

> is FastingBS is really playing a good role in RestingBP ?? 

```{r}
Bs_modelBP <- lm(RestingBP ~ FastingBS, data = subset)
summary(Bs_modelBP)
```


#### Logistic Regresion

> Does age play significant role in HeartDisease

```{r}
# glm generalized linear regression
glm_reg <- glm(HeartDisease ~ Age, data = heart_dataset, family = "binomial")
glm_reg
```

```{r}
range(heart_dataset$Age)

x_axis <- seq(min(range(heart_dataset$Age)), max(range(heart_dataset$Age)),3)
x_axis
y_axis <- predict(glm_reg, list(Age=x_axis),type = "response")
y_axis

plot(heart_dataset$Age,heart_dataset$HeartDisease)
plot(x_axis,y_axis)
lines(x_axis,y_axis)
```

###     

#### Splitting Dataset into Training and Test

```{r}

set.seed(17)
heart_train_data_index<- sample(1:nrow(heart_dataset),0.6*nrow(heart_dataset))

heart_training.set <- heart_dataset[heart_train_data_index,-c(2,7,9,11)]
head(heart_training.set)
heart_test.set <- heart_dataset[-heart_train_data_index,-c(2,7,9,11)]
head(heart_test.set)

```
 
#### KNN model on Chest Pain

```{r}
knn_indep_train_set =heart_training.set[-2] # remove target variable from dataset
knn_indep_train_set

knn_indep_test_set = heart_test.set[-2]
knn_indep_test_set

knn_target_train_set = heart_training.set$ChestPainType
head(knn_target_train_set)

knn_target_test_set = heart_test.set$ChestPainType
head(knn_target_test_set)


knn_pred_target_test_set = knn(knn_indep_train_set,knn_indep_test_set, knn_target_train_set,k=5)
knn_pred_target_test_set

cnfMtrx <- table(knn_pred_target_test_set,knn_target_test_set)
cnfMtrx

plot_confusion_matrix(as_tibble(cnfMtrx),target_col = "knn_target_test_set", 
                      prediction_col = "knn_pred_target_test_set",
                      counts_col = "n")
chestPainAccuracy <- sum(diag(cnfMtrx))/sum(cnfMtrx)
chestPainAccuracy

```


#### Decision Tree model on Chest Pain

>run decision tree model on training set

```{r}
#Error in trafo(data = data, numeric_trafo = numeric_trafo, factor_trafo = factor_trafo, : data class “character” is not supported

heart_training.set$ChestPainType <- as.factor(heart_training.set$ChestPainType)
head(heart_training.set)
heart_training_ctree_model <- ctree(ChestPainType ~ ., data = heart_training.set)
heart_training_ctree_model
plot(heart_training_ctree_model, type="simple")
```

> prediction on test set

```{r}
heart_ctree_prediction <- predict(heart_training_ctree_model,heart_test.set)
heart_ctree_prediction
```

```{r}
ctreeMtrx <- table(heart_ctree_prediction,heart_test.set$ChestPainType)
ctreeMtrx

ctreeAccuracy <- sum(diag(ctreeMtrx))/sum(ctreeMtrx)
ctreeAccuracy 

```
#### Random forest on Chest Pain

```{r}

random_forest_chestPain<-randomForest(ChestPainType~.,data=heart_training.set)
plot(random_forest_chestPain)
print(random_forest_chestPain)


print(importance(random_forest_chestPain,type = 2))

heartdataset_randomforest_prediction<-predict(random_forest_chestPain,heart_test.set)
heartdataset_randomforest_prediction

random_forest_Mtrx_chestPain <- table(heartdataset_randomforest_prediction,heart_test.set$ChestPainType)
random_forest_Mtrx_chestPain
random_forest_accuracy_chestPain <- sum(diag(random_forest_Mtrx_chestPain))/sum(random_forest_Mtrx_chestPain)
random_forest_accuracy_chestPain

```

#### Model Accuracy comparision for chestpain Type

```{r}
Model <- c('KNN','Decision Tree','Random Forest')
Accuracy <- c(chestPainAccuracy, ctreeAccuracy , random_forest_accuracy_chestPain)
ChestPain_accuracy_table <- data.frame(Model, Accuracy)
ChestPain_accuracy_table
```


#### KNN model on Gender

```{r}
#Predict Gender
set.seed(17)
heart_train_data_index1<- sample(1:nrow(heart_dataset),0.6*nrow(heart_dataset))
head(heart_train_data_index1)

heart_training.set1 <- heart_dataset[heart_train_data_index1,-c(3,7,9,11)]
head(heart_training.set1)
heart_test.set1 <- heart_dataset[-heart_train_data_index1,-c(3,7,9,11)]
head(heart_test.set1)

knn_indep_train_set1 =heart_training.set1[-2] # remove target variable from dataset
knn_indep_train_set1

knn_indep_test_set1 = heart_test.set1[-2]
knn_indep_test_set1

knn_target_train_set1 = heart_training.set1$Sex
head(knn_target_train_set1)

knn_target_test_set1 = heart_test.set1$Sex
head(knn_target_test_set1)

knn_pred_target_test_set1 = knn(knn_indep_train_set1,knn_indep_test_set1, knn_target_train_set1,k=3)
knn_pred_target_test_set1

mtrx <- table(knn_pred_target_test_set1,knn_target_test_set1)
mtrx

plot_confusion_matrix(as_tibble(mtrx),target_col = "knn_target_test_set1", 
                      prediction_col = "knn_pred_target_test_set1",
                      counts_col = "n")

genderAccuracy <- sum(diag(mtrx))/sum(mtrx)
genderAccuracy
```


#### Decision tree model on gender

```{r}
heart_training.set1$Sex <- as.factor(heart_training.set1$Sex)
#head(heart_training.set)
heart_training_ctree_model_gender <- ctree(Sex ~ ., data = heart_training.set1)
heart_training_ctree_model_gender
plot(heart_training_ctree_model_gender, type="simple")

# Prediction on test set
heart_ctree_prediction_gender <- predict(heart_training_ctree_model_gender,heart_test.set1)
heart_ctree_prediction_gender

ctreeMtrxGender <- table(heart_ctree_prediction_gender,heart_test.set1$Sex)
ctreeMtrxGender
ctreeAccuracyGender <- sum(diag(ctreeMtrxGender))/sum(ctreeMtrxGender)
ctreeAccuracyGender 
```

#### Random forest in Gender

```{r}

random_forest_gender<-randomForest(Sex~.,data=heart_training.set1)
plot(random_forest_gender)
print(random_forest_gender)

print(importance(random_forest_gender,type = 2))

heartdataset_randomforest_prediction_gender<-predict(random_forest_gender,heart_test.set1)
heartdataset_randomforest_prediction_gender

random_forest_Mtrx_gender <- table(heartdataset_randomforest_prediction_gender,heart_test.set1$Sex)
random_forest_Mtrx_gender
random_forest_accuracy_gender <- sum(diag(random_forest_Mtrx_gender))/sum(random_forest_Mtrx_gender)
random_forest_accuracy_gender 
```

#### Model Accuracy comparision for chestpain Type

```{r}
Model <- c('KNN','Decision Tree','Random Forest')
Accuracy <- c(genderAccuracy, ctreeAccuracyGender , random_forest_accuracy_gender)
Gender_accuracy_table <- data.frame(Model, Accuracy)
Gender_accuracy_table
```

